---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Elayna Seago"
date: "2/16/22"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1
There are 3 dennys in Alaska

```{r num-dennys-alaska}

dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

```


### Exercise 2
There are 2 laquintas
```{r num-laquinta-alaska}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```


### Exercise 3

```{r join-data}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```


### Exercise 4
There are 6 observations and the variables are address, city, zip, long and lat for each dennys and laquinta location, and state.

```{r num-obs-and-variables}
nrow(dn_lq_ak)
ls(dn_lq_ak)
```



### Exercise 5

```{r make-haversine-func}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```


### Exercise 6

```{r calc-dist}
dn_lq_ak <- mutate(dn_lq_ak, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

```

### Exercise 7

```{r find-min-dis}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_ak_mindist

```

### Exercise 8
Because there are only 3 Dennys and 2 laquintas in Alaska, the distribution is not particularly informative. Dennys and lquintas are close together, on average they are 4.41 miles apart.

```{r}
mean(dn_lq_ak_mindist$closest)

ggplot(dn_lq_ak_mindist, aes(x = closest))+
  geom_histogram() +
  geom_vline(xintercept = mean(dn_lq_ak_mindist$closest), lwd = 2 , linetype = 2 , color = "black")
```


### Exercise 9
NC
There is a slight positive skew, but noting the change in X axis, the distance between laquintas and Dennys is much larger in NC than in AK.The mean distance to the closest is 65 miles.
```{r NC-num-join}

dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)

lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)

dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc

```

```{r NC-compute}

dn_lq_nc <- mutate(dn_lq_nc, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_nc_mindist

mean(dn_lq_nc_mindist$closest)

ggplot(dn_lq_nc_mindist, aes(x = closest))+
  geom_histogram() +
  geom_vline(xintercept = mean(dn_lq_nc_mindist$closest), lwd = 2 , linetype = 2 , color = "black")
```

### Exercise 10
TX
The mean distance to the closest Dennys is 5 miles. The distribution is very positively skewed. This state is the one where the joke seems to make the most sense (the mean is smaller aka closer for AK, but the sample size is very small).
```{r tx-num-join}

dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)

lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)

dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx
```

```{r tx-compute}

dn_lq_tx <- mutate(dn_lq_tx, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_tx_mindist

mean(dn_lq_tx_mindist$closest)

ggplot(dn_lq_tx_mindist, aes(x = closest))+
  geom_histogram() +
  geom_vline(xintercept = mean(dn_lq_tx_mindist$closest), lwd = 2 , linetype = 2 , color = "black")
```

### Exercise 11
SC
On average, dennys and laquintas are 32 miles apart in SC. This distribution is wide and flat and not bell curved at all.
```{r}
dn_sc <- dennys %>%
  filter(state == "SC")
nrow(dn_sc)

lq_sc <- laquinta %>%
  filter(state == "SC")
nrow(lq_sc)

dn_lq_sc <- full_join(dn_sc, lq_sc, by = "state")
dn_lq_sc
```

```{r}
dn_lq_sc <- mutate(dn_lq_sc, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

dn_lq_sc_mindist <- dn_lq_sc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_sc_mindist
mean(dn_lq_sc_mindist$closest)

ggplot(dn_lq_sc_mindist, aes(x = closest))+
  geom_histogram() +
  geom_vline(xintercept = mean(dn_lq_sc_mindist$closest), lwd = 2 , linetype = 2 , color = "black")
```

